---
title: "benchmarks"
output: html_document
---

# Benchmarks for IREGNET

Loading the required packges
```{r loading}
library(iregnetoptim)
library(iregnet)
library(glmnet)
library(survival)
library(directlabels)
library(ggplot2)
library(microbenchmark)
library(dplyr) 
```


## Uncensored Data

- **Prostate Dataset**

**Packages: Iregnet, Iregnet Optimised, Glmnet**
```{r test}
data("Prostate", package = "lasso2")
X = as.matrix(Prostate[, c(2:9)])
Y = matrix(c(Prostate[, 1],Prostate[, 1]), nrow = nrow(Prostate), ncol = 2)
colnames(Y)[c(1,2)] <- "lcalvol"
res <- data.frame() #Result data frame
for(i in c(4:(nrow(X)/5))*5)
{
  evaltime <- microbenchmark(iregnet(X[1:i,], Y[1:i,]), 
                             iregnetoptim(X[1:i,], Y[1:i,]), 
                             glmnet(X[1:i,], Y[1:i,1]), 
                             times = 100L)
  res <- bind_rows(res, data.frame(i, list(summary(evaltime)[,c('lq','mean','uq')])))
}
res <- cbind.data.frame(c("IREGNET", "Iregnet Optimization", "GLMNET"), res)
names(res) <- c("expr", names(res)[2:5])
p <- ggplot(res, aes(x = i))+
  geom_ribbon(aes(ymin = lq, ymax = uq, fill = expr, group = expr), alpha = 1/2)+
  geom_line(aes(y = mean, group = expr, colour = expr))+
  ggtitle('Runtime(in milliseconds) vs Dataset Size') +
  xlab('Dataset Size') +
  ylab('Runtime (in milliseconds)')
direct.label(p,"angled.boxes")
```

A weird thing, did not fit for many values of i
```{r weird}
data("Prostate", package = "lasso2")
X = as.matrix(Prostate[, c(2:9)])
Y = matrix(c(Prostate[, 1],Prostate[, 1]), nrow = nrow(Prostate), ncol = 2)
colnames(Y)[c(1,2)] <- "lcalvol"
res <- data.frame() #Result data frame
for(i in c(30:nrow(X)))
{
  X_copy <- X
  Y_copy <- Y
  evaltime <- microbenchmark(iregnet(X[1:i,], Y[1:i,]), 
                             iregnetoptim(X_copy[1:i,], Y_copy[1:i,]), 
                             glmnet(X[1:i,], Y[1:i,1]), 
                             times = 100L)
  res <- bind_rows(res, data.frame(i, list(summary(evaltime)[,c('lq','mean','uq')])))
}
res <- cbind.data.frame(c("IREGNET", "Iregnet Optimization", "GLMNET"), res)
names(res) <- c("expr", names(res)[2:5])
p <- ggplot(res, aes(x = i))+
  geom_ribbon(aes(ymin = lq, ymax = uq, fill = expr, group = expr), alpha = 1/2)+
  geom_line(aes(y = mean, group = expr, colour = expr))+
  ggtitle('Runtime(in milliseconds) vs Dataset Size') +
  xlab('Dataset Size') +
  ylab('Runtime (in milliseconds)')
direct.label(p,"angled.boxes")
```
Using the lambda path from glmnet
```{r prostate_lambda}
data("Prostate", package = "lasso2")
X = as.matrix(Prostate[, c(2:9)])
Y = matrix(c(Prostate[, 1],Prostate[, 1]), nrow = nrow(Prostate), ncol = 2)
colnames(Y)[c(1,2)] <- "lcalvol"
#Centering the data
Y <- apply(Y, 2, function(y) y - mean(y))
X <- apply(X, 2, function(x) x - mean(x))
res <- data.frame() #Result data frame
for(i in 20:nrow(X))
{
  glm <- glmnet(X[1:i,], Y[1:i,1])
  evaltime <- microbenchmark(iregnet(X[1:i,], Y[1:i,], 
                                     num_lambda = length(glm$lambda), lambda = glm$lambda), 
                             iregnetoptim(x <- X[1:i,],y <- Y[1:i,], 
                                     num_lambda = length(glm$lambda), lambda = glm$lambda),
                             glmnet(X[1:i,], Y[1:i,1]), 
                             times = 100L)
  res <- bind_rows(res, data.frame(i, list(summary(evaltime)[,c('lq','mean','uq')])))
}
res <- cbind.data.frame(c("IREGNET", "IREGNET Optimized", "GLMNET"), res)
names(res) <- c("expr", names(res)[2:5])
p <- ggplot(res, aes(x = i))+
  geom_ribbon(aes(ymin = lq, ymax = uq, fill = expr, group = expr), alpha = 1/2)+
  geom_line(aes(y = mean, group = expr, colour = expr))+
  ggtitle('Runtime(in milliseconds) vs Dataset Size') +
  xlab('Dataset Size') +
  ylab('Runtime (in milliseconds)')
direct.label(p,"angled.boxes")
```


- **Random Dataset for 1000 observations**
**With packages glmnet, iregnet and iregnet optimised**
```{r rand_1000}
X <- rnorm(5000, 1, 1.5) %>% matrix(nrow = 1000, ncol = 5)
Y <- rnorm(1000, 1, 1.5) %>% matrix(nrow = 1000, ncol = 1)
Y = matrix(c(Y, Y), nrow = 1000, ncol = 2)
res <- data.frame() #Result data frame
for(i in c(1:50)*20)
{
  glm <- glmnet(X[1:i,], Y[1:i,1])
  evaltime <- microbenchmark(iregnet(X[1:i,], Y[1:i,], 
                                     num_lambda = length(glm$lambda), lambda = glm$lambda), 
                             iregnetoptim(X[1:i,], Y[1:i,], 
                                          num_lambda = length(glm$lambda), lambda = glm$lambda),
                             glmnet(X[1:i,], Y[1:i,1]), 
                             times = 100L)
  res <- bind_rows(res, data.frame(i, list(summary(evaltime)[,c('lq','mean','uq')])))
}
res <- cbind.data.frame(c("IREGNET", "IREGNET Optimized", "GLMNET"), res)
names(res) <- c("expr", names(res)[2:5])

res_temp <- res[34:150,]

p <- ggplot(res, aes(x = i))+
  geom_ribbon(aes(ymin = lq, ymax = uq, fill = expr, group = expr), alpha = 1/2)+
  geom_line(aes(y = mean, group = expr, colour = expr))+
  ggtitle('Runtime(in milliseconds) vs Dataset Size') +
  xlab('Dataset Size') +
  ylab('Runtime (in milliseconds)')
direct.label(p,"angled.boxes")

p_temp <- ggplot(res_temp, aes(x = i))+
  geom_ribbon(aes(ymin = lq, ymax = uq, fill = expr, group = expr), alpha = 1/2)+
  geom_line(aes(y = mean, group = expr, colour = expr))+
  ggtitle('Runtime(in milliseconds) vs Dataset Size') +
  xlab('Dataset Size') +
  ylab('Runtime (in milliseconds)')
direct.label(p_temp,"angled.boxes")
```

**With packages iregnet and glmnet**
```{r rand_1000_non_optimised}
X <- rnorm(5000, 1, 1.5) %>% matrix(nrow = 1000, ncol = 5)
Y <- rnorm(1000, 1, 1.5) %>% matrix(nrow = 1000, ncol = 1)
Y = matrix(c(Y, Y), nrow = 1000, ncol = 2)
res <- data.frame() #Result data frame
for(i in c(1:50)*20)
{
  glm <- glmnet(X[1:i,], Y[1:i,1])
  evaltime <- microbenchmark(iregnet(X[1:i,], Y[1:i,], 
                                     num_lambda = length(glm$lambda), lambda = glm$lambda), 
                             glmnet(X[1:i,], Y[1:i,1]), 
                             times = 100L)
  res <- bind_rows(res, data.frame(i, list(summary(evaltime)[,c('lq','mean','uq')])))
}
res <- cbind.data.frame(c("IREGNET", "GLMNET"), res)
names(res) <- c("expr", names(res)[2:5])
p <- ggplot(res, aes(x = i))+
  geom_ribbon(aes(ymin = lq, ymax = uq, fill = expr, group = expr), alpha = 1/2)+
  geom_line(aes(y = mean, group = expr, colour = expr))+
  ggtitle('Runtime(in milliseconds) vs Dataset Size') +
  xlab('Dataset Size') +
  ylab('Runtime (in milliseconds)')
direct.label(p,"angled.boxes")
```




- **Random Dataset for 100000 observations**

**Packages: Iregnet, Iregnet Optimised, Glmnet**
```{r random}
## Randomly generated data
X <- rnorm(500000, 1, 1.5) %>% matrix(nrow = 100000, ncol = 5)
Y <- rnorm(100000, 1, 1.5) %>% matrix(nrow = 100000, ncol = 1)
Y = matrix(c(Y, Y), nrow = 100000, ncol = 2)
res <- data.frame() 
for(i in c(1:20)*5000)
{
  evaltime <- microbenchmark(iregnet(X[1:i,], Y[1:i,]), 
                             iregnetoptim(X[1:i,], Y[1:i,]), 
                             glmnet(X[1:i,], Y[1:i,1]), 
                             times = 10L)
  res <- bind_rows(res, data.frame(i, list(summary(evaltime)[,c('lq','mean','uq')])))
}
res <- cbind.data.frame(c("IREGNET", "Iregnet Optimization", "GLMNET"), res)
names(res) <- c("expr", names(res)[2:5])
p <- ggplot(res, aes(x = i))+
  geom_ribbon(aes(ymin = lq, ymax = uq, fill = expr, group = expr), alpha = 1/2)+
  geom_line(aes(y = mean, group = expr, colour = expr))+
  ggtitle('Runtime(in milliseconds) vs Dataset Size') +
  xlab('Dataset Size') +
  ylab('Runtime (in milliseconds)')
direct.label(p,"angled.boxes")
```

## Right Censored Data

- **Ovarian Dataset**

**Packages: Iregnet, Iregnet Optimised, Survreg**
**Unregularised Solution**
```{r ovar_unreg}
data("ovarian")
X <- cbind(ovarian$ecog.ps, ovarian$rx)
res <- data.frame() #Result data frame
for(i in c(20:nrow(X)))
{
  evaltime <- microbenchmark(iregnet(X[1:i,], 
                                     Surv(ovarian$futime[1:i], ovarian$fustat[1:i]),
                                     num_lambda = 1, lambda = 0), 
                             iregnetoptim(X[1:i,], 
                                          Surv(ovarian$futime[1:i], ovarian$fustat[1:i]),
                                          num_lambda = 1, lambda = 0),
                             survreg(Surv(futime, fustat) ~ X[1:i,], data = ovarian[1:i,]), 
                             times = 100L)
  res <- bind_rows(res, data.frame(i, list(summary(evaltime)[,c('lq','mean','uq')])))
}
res <- cbind.data.frame(c("IREGNET", "Iregnet Optimization", "Survreg"), res)
names(res) <- c("expr", names(res)[2:5])
p <- ggplot(res, aes(x = i))+
  geom_ribbon(aes(ymin = lq, ymax = uq, fill = expr, group = expr), alpha = 1/2)+
  geom_line(aes(y = mean, group = expr, colour = expr))+
  ggtitle('Runtime(in milliseconds) vs Dataset Size') +
  xlab('Dataset Size') +
  ylab('Runtime (in milliseconds)')
direct.label(p,"angled.boxes")
```
**Regularised Solution with 100 lambdas**
```{r ovar}
data("ovarian")
X <- cbind(ovarian$ecog.ps, ovarian$rx)
res <- data.frame() #Result data frame
for(i in c(15:nrow(X)))
{
  evaltime <- microbenchmark(iregnet(X[1:i,], 
                                     Surv(ovarian$futime[1:i], ovarian$fustat[1:i])), 
                             iregnetoptim(X[1:i,], 
                                          Surv(ovarian$futime[1:i], ovarian$fustat[1:i])),
                             survreg(Surv(futime, fustat) ~ X[1:i,], data = ovarian[1:i,]), 
                             times = 100L)
  res <- bind_rows(res, data.frame(i, list(summary(evaltime)[,c('lq','mean','uq')])))
}
res <- cbind.data.frame(c("IREGNET", "Iregnet Optimization", "Survreg"), res)
names(res) <- c("expr", names(res)[2:5])
p <- ggplot(res, aes(x = i))+
  geom_ribbon(aes(ymin = lq, ymax = uq, fill = expr, group = expr), alpha = 1/2)+
  geom_line(aes(y = mean, group = expr, colour = expr))+
  ggtitle('Runtime(in milliseconds) vs Dataset Size') +
  xlab('Dataset Size') +
  ylab('Runtime (in milliseconds)')
direct.label(p,"angled.boxes")
```

## Interval Censored Data

**Neuroblastoma**
**Packages: Iregnet, Iregnet Optimised, Glmnet**
```{r interval, eval = FALSE}
data("neuroblastomaProcessed", package = "penaltyLearning")
X <- neuroblastomaProcessed$feature.mat
Y <- neuroblastomaProcessed$target.mat
X = X[,apply(X,2,function(x){
  return(var(x)!=0)
  })]
res <- data.frame() #Result data frame
# for(i in c((6:12)*250, 3418)
for(i in c(6:12)*250)
{
  evaltime <- microbenchmark(iregnet(X[1:i,], Y[1:i,]), 
                             iregnetoptim(X[1:i,], Y[1:i,]), 
                             times = 10L)
  res <- bind_rows(res, data.frame(i, list(summary(evaltime)[,c('lq','mean','uq')])))
}
res <- cbind.data.frame(c("IREGNET", "Iregnet Optimization", "Survreg"), res)
names(res) <- c("expr", names(res)[2:5])
p <- ggplot(res, aes(x = i))+
  geom_ribbon(aes(ymin = lq, ymax = uq, fill = expr, group = expr), alpha = 1/2)+
  geom_line(aes(y = mean, group = expr, colour = expr))+
  ggtitle('Runtime(in milliseconds) vs Dataset Size') +
  xlab('Dataset Size') +
  ylab('Runtime (in milliseconds)')
direct.label(p,"angled.boxes")
```